$R^*$: A robust MCMC convergence diagnostic with uncertainty using gradient-boosted machines

### 2020-09-04 src/s_eight_schools.Rmd
addressing andrew gelman's comments including transposing fig 12

# Plot R* versus quantiles
```{r}
x <- rstan::extract(fit_cp, permuted=F)
g1 <- plot_r_star_quantiles(1, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 1.7), breaks = c(0.6, 0.8, 1, 1.2, 1.4, 1.6))

g2 <- plot_r_star_quantiles(2, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 1.7), breaks = c(0.6, 0.8, 1, 1.2, 1.4, 1.6))

library(ggExtra)
library(gridExtra)
f <- ecdf(x[, , 2])
p <- x[, , 2] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none",
        axis.title.x = element_blank()) +
  xlab("Iteration") +
  ylab("Quantile") +
  ylim(0, 1) +
  coord_flip()
ga <- ggMarginal(p, type="density", margins="x", groupColour = T,
                 xparams = list(bw=0.05))

f <- ecdf(x[, , 1])
p <- x[, , 1] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none",
        axis.title.x = element_blank()) +
  xlab("Iteration") +
  ylab("Quantile") +
  ylim(0, 1) +
  coord_flip()
gb <- ggMarginal(p, type="density", margins="x", groupColour = T,
                 xparams = list(bw=0.05))

@@

pdf("../output/eight_schools_r_star_quantiles.pdf", width = 6, height = 6)
plot_grid(arrangeGrob(gb, top="A. mu"), arrangeGrob(ga, top="B. tau"), g1, g2,nrow = 2, rel_heights = c(1/2, 1/4))
dev.off()
```

>>>>

# Plot R* versus quantiles
```{r}
x <- rstan::extract(fit_cp, permuted=F)
g1 <- plot_r_star_quantiles(1, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 1.7), breaks = c(0.6, 0.8, 1, 1.2, 1.4, 1.6)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip()

g2 <- plot_r_star_quantiles(2, x, nsim=40, split_chains = F) +
  scale_y_continuous(limits = c(0.6, 1.7), breaks = c(0.6, 0.8, 1, 1.2, 1.4, 1.6)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip()

library(ggExtra)
library(gridExtra)
f <- ecdf(as.vector(x[, , 2]))
p <- x[, , 2] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none") +
  xlab("Iteration") +
  ylab(TeX("B. $\\tau$ quantile")) +
  ylim(0, 1)
ga <- ggMarginal(p, type="density", margins="y", groupColour = T,
                 yparams = list(bw=0.05))

f <- ecdf(as.vector(x[, , 1]))
p <- x[, , 1] %>% 
  melt(id.vars="NULL") %>% 
  mutate(value=f(value)) %>% 
  ggplot(aes(x=iterations, y=value, colour=as.factor(chains))) +
  geom_line(alpha=0.8) +
  geom_point(aes(size=NA)) +
  scale_size_identity() +
  scale_color_brewer(palette = "Dark2") +
  theme(text = element_text(colour="black", size=11),
          axis.text = element_text(colour="black", size=10),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  xlab("Iteration") +
  ylab(TeX("A. $\\mu$ quantile")) +
  ylim(0, 1)
gb <- ggMarginal(p, type="density", margins="y", groupColour = T,
                 yparams = list(bw=0.05))

@@

pdf("../output/eight_schools_r_star_quantiles1.pdf", width = 6, height = 4)
plot_grid(gb, g1, ga, g2, nrow = 2, rel_widths = c(1, 0.4), rel_heights = c(0.9, 1))
dev.off()
```

### 2020-09-04 write_up/arxiv_submission_aug_2020/ms_arxiv_2020.tex
addressing andrew gelman's comments including transposing fig 12

The predominant diagnostic method for determining whether practical convergence has occurred relies on the fact that the posterior distribution is the unique stationary distribution for an MCMC sampler. Therefore, it would appear that, if an MCMC sampling distribution stops changing, then convergence has occurred. Unfortunately, anyone who uses MCMC knows that it is full of false dawns: chains can easily become stuck in areas of parameter space, and observation over short intervals mean the sampling distribution \textit{appears} converged. Like furious bees trapped in a room of a house \citep{lambertbees}, MCMC samplers may fail to move due to the narrow gaps that join neighbouring areas. With MCMC, absence of evidence of new areas of high posterior density is, time and again, not evidence of their absence.

@@

Here, we introduce $R^*$, a new convergence metric. This statistic is built on the intuition that, if chains are mixed, it should not be possible to discern from a draw's value the chain that generated it. Rephrased, it should not be possible to predict the chain that \textit{caused} a draw. In this vein, we use a machine learning (ML) classifier to measure convergence. Specifically, we train a ML classifier to predict the chain that generated each observation. By evaluating the performance of the classifier on a held-out test set, this provides a new convergence metric. To maximise predictive accuracy, our chosen ML classifier naturally exploits differences in the full joint distributions between chains, which means it's sensitive to variations across the joint distribution of target model dimensions unlike most existent convergence diagnostics. Our statistic, unlike its $\widehat{R}$ cousins, is scalar valued for multivariate distributions: one model provides a single $R^*$, whereas $\widehat{R}$ has separate values for each univariate marginal distribution. However, the ML classifier we use can straightforwardly be interrogated to determine which parameters were most important for generating predictive accuracy. For our ML classifier, we use gradient-boosted regression trees \citep{friedman2001greedy,greenwell2019package} (``GBM''), since these are known to perform well for the types of tabular data that our problem presents \citep{chollet2018}. For the types of problem we test, $R^*$ calculation is of a speed comparable to some of the newer $\widehat{R}$ measures calculated (typically $\mathcal{O}$(seconds) to calculate), although for models with 10,000s of parameters and many iterations, the time taken is longer. It is also insensitive to GBM's hyperparameters and provides a measure of convergence robust to various Markov chain pathologies. In addition, since GBMs can output predicted class probabilities, we obtain uncertainty measures for $R^*$, which we find provides a useful summary of MCMC convergence. $R^*$ can straightforwardly be incorporated into existing software libraries to provide a complementary convergence metric alongside more established measures.

@@

\section{Method}\label{sec:method}
If Markov chains have not mixed, it is possible to determine to which chain a draw belongs from its value. This is possible if there are differences in the sampling distribution for any dimension, $\theta$, in the target distribution (Fig. \ref{fig:marginal}): in this case, if the marginal distributions differ between chains, this information can be used to predict which chain a draw belongs to. It  is also possible to predict the chain that generated a given draw if there are differences in the joint distribution of two (or more) dimensions of the target, even if the marginal distributions are the same (Fig. \ref{fig:joint}).

@@

\subsection{Heterogeneity in chain variance: autoregressive example}\label{sec:heterogeneity}
In this section, we show how $R^*$ is able to detect heterogeneous variance across Markov chains. Apart from demonstrating how $R^*$ performs, this section also investigates the sensitivity of this measure: across different training and testing sets (\S\ref{sec:heterogeneity_stochasticity}) and different draws from the ML model predicted probability simplex (\S\ref{sec:heterogeneity_uncertainty}); to differing numbers of chains (\S\ref{sec:heterogeneity_numchains}); and, finally, to the ML model hyperparameters (\S\ref{sec:heterogeneity_robustness}). The experiments we use to study these issues are all of similar form to the following data generating process: four Markov chains are generated, where each samples from an autoregressive order 1 (AR(1)) process of the form,
%
\begin{equation}\label{eq:ar1}
X_t = \rho X_{t-1} + \epsilon_t,
\end{equation}
%
where $\epsilon_t\stackrel{i.i.d.}{\sim}\mathcal{N}(0, \sigma)$, $\rho=0.3$ and $t=1,2,...,2000$. Three of the chains share the same $\sigma=1$, whereas the other chain has $\sigma=1/3$, so that it has $1/3$ of the (unconditional) standard deviation of the others.

@@

Rather than run the MCMC sampler for more iterations, we move to a ``non-centered'' parameterisation, which introduces auxillary variables $\boldsymbol{z}\in\mathbb{R}^{250}$ that don't affect $p(\boldsymbol{x})$ but facilitate sampling from it. This model has the form,
%
\begin{align}
\boldsymbol{A}^{-1} = \boldsymbol{L}\boldsymbol{L}^T,\qquad
\boldsymbol{x} = \boldsymbol{L} \boldsymbol{z},\qquad
z_j\sim \mathcal{N}(0, 1), \text{ for } j = 1,2,...,250.
\end{align}
%
where $\boldsymbol{L}$ is the Cholesky decomposition of the covariance matrix, $\boldsymbol{A}^{-1}$. Fig. \ref{fig:mvt}A shows the $R^*$ distribution resultant from 10,000 NUTS iterations in this case: now the distribution has mean $R^*=1.00$. Fig. \ref{fig:mvt}B shows the $\widehat{R}$ values for each $x$ parameter in this model, and, echoing the result for $R^*$, $\widehat{R}<1.01$ in all cases; further, bulk- and tail-$\text{ESS}>400$ for all parameters.

@@

In addition, we also use Stan to sample from an ``alternative'' parameterisation of the Cauchy, based on a scale mixture of Gaussians \citep{vehtari2019rank},
%
\begin{align}
a_j \sim  \N(0,1), \qquad
b_j \sim  \text{Gamma}(0.5, 0.5), \qquad
x_j =  a_j/\sqrt{b_j}.
\end{align}
%
The distribution of the $x$ vector is the same under both parameterisations, although the thin-tailed $(a,b)$ vectors define a higher dimensional posterior that improves sampling efficiency.

@@

Comparing our ``omniscient'' measure of convergence shown in Fig. \ref{fig:cauchy_convergence}A, with the various derived measures, all show a similar pattern: as sample size increases, the various statistics tend towards convergence. The rate at which these converge differs though, and $R^*$ (Fig. \ref{fig:cauchy_convergence}D) appears at least, qualitatively, most similar to our ``omniscient'' measure.

@@

The model can be parameterised in two ways, as described in \cite{vehtari2019rank}. The simplest way is referred to as the ``centered'' parameterisation and exactly mirrors the underlying statistical model,
%
\begin{align*}
\theta_j &\sim \N(\mu, \tau), \\
y_j &\sim \N(\theta_j, \sigma_j).
\end{align*}
%
The ``non-centered'' parameterisation recodes this model in a way that does not affect the joint distribution of $(\theta, \mu, \tau, \sigma)$ but makes it easier to sample from it, by introducing auxillary variables, $\tilde \theta_j$. This can be written as,
%
\begin{align*}
\tilde{\theta}_j &\sim \N(0, 1), \\
\theta_j &= \mu + \tau \tilde{\theta}_j,\\
y_j &\sim \N(\theta_j, \sigma_j).
\end{align*}
%
In both cases, $\theta_j$ are the treatment effects in the eight schools, and $(\mu, \tau)$ represent the population mean and standard deviation 
of the distribution of these effects. In the centered parameterization, the $\theta_j$ are parameters, whereas in the non-centered parameterization, the $\tilde{\theta}_j$ are parameters and $\theta_j$ is a derived quantity.

@@

\subsubsection{Understanding chain classification}
To probe the predictive power of the ML classifier, we investigated how predictive accuracy varies across parameter space. After fitting the GBM model, we group MCMC draws into deciles and draw from the $R^*$ distribution for each decile. In Fig. \ref{fig:eight_schools_r_star_quantiles}, we show the results of this exercise for (A) $\mu$ and (B) $\tau$. In the top row of this figure, we show the path of four MCMC chains (here we did not split chains when calculating $R^*$ to simplify visualisations) across the quantiles of each parameter space. Above the top panel, we show the marginal distributions for each chain. In the bottom row, we show 40 $R^*$ draws for each decile, which were generated according to Algorithm \ref{alg:R_star_uncertainty} using a GBM fit to all draws. In essence, the top rows explain the variation in $R^*$ in the bottom panels: if chains become stuck in regions of parameter space, this causes differences between the marginal distributions of the chains; these differences, in turn, allow a ML model to predict the generative chain in those same sticky regions. For example, for $\mu$, the purple chain became stuck around the middle quantile, forcing a difference in its marginal distribution in that region, which resulted in $R^*>1$ for the corresponding decile. Similarly, for $\tau$, the purple chain became stuck in the lowest quantiles, elevating its marginal distribution there and resulting in improved ML predictive accuracy.

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{eight_schools_r_star_quantiles.pdf}}
	\caption{\textbf{Eight schools example: quantile $R^*$ plots.} Column A shows plots for $\mu$; Column B for $\tau$. In each column, we show the path of the four individual chains above and 40 $R^*$ draws obtained using Algorithm \ref{alg:R_star_uncertainty} for each parameter quantile below. Above the top row, we show the marginal distribution of each chain estimated via kernel density estimation using Gaussian kernels. Note that, in the bottom plots, jitter has been added to the data points.}
	\label{fig:eight_schools_r_star_quantiles}
\end{figure}

@@

In \S\ref{sec:prostate}, we fit Bayesian models with many 1000s of parameters then used $R^*$ to diagnose convergence, finding that $R^*$ was considerably more expensive to calculate than $\widehat{R}$. The computational cost of training GBMs has been suggested to be $\mathcal{O}(N k)$, where $N$ is the number of training data points and $k$ is the number of dimensions of the target \citep{complexity2018}. If so, this suggests that larger models (usually needing more MCMC iterations) may currently be beyond the reach of $R^*$.

@@

>>>>

The predominant diagnostic method for determining whether practical convergence has occurred relies on the fact that the posterior distribution is the unique stationary distribution for an MCMC sampler. Therefore, it would appear that, if an MCMC sampling distribution stops changing, then convergence has occurred. Unfortunately, anyone who uses MCMC knows that it is full of false dawns: chains can easily become stuck in areas of parameter space, and observation over short intervals mean the sampling distribution \textit{appears} converged \citep{gelman1992single}. Like furious bees trapped in a room of a house \citep{lambertbees}, MCMC samplers may fail to move due to the narrow gaps that join neighbouring areas. With MCMC, absence of evidence of new areas of high posterior density is, time and again, not evidence of their absence.

@@

Here, we introduce $R^*$, a new convergence metric. This statistic is built on the intuition that, if chains are mixed, it should not be possible to discern from a draw's value the chain that generated it. Rephrased, it should not be possible to predict which draws come from which chain. In this vein, we use a machine learning (ML) classifier to measure convergence. Specifically, we train a ML classifier to predict the chain that generated each observation. By evaluating the performance of the classifier on a held-out test set, this provides a new convergence metric. To maximise predictive accuracy, our chosen ML classifier naturally exploits differences in the full joint distributions between chains, which means it's sensitive to variations across the joint distribution of target model dimensions unlike most existent convergence diagnostics. Our statistic, unlike its $\widehat{R}$ cousins, is scalar valued for multivariate distributions: one model provides a single $R^*$, whereas $\widehat{R}$ has separate values for each univariate marginal distribution. However, the ML classifier we use can straightforwardly be interrogated to determine which parameters were most important for generating predictive accuracy. For our ML classifier, we use gradient-boosted regression trees \citep{friedman2001greedy,greenwell2019package} (``GBM''), since these are known to perform well for the types of tabular data that our problem presents \citep{chollet2018}. For the types of problem we test, $R^*$ calculation is of a speed comparable to some of the newer $\widehat{R}$ measures calculated (typically $\mathcal{O}$(seconds) to calculate), although for models with 10,000s of parameters and many iterations, the time taken is longer. It is also insensitive to GBM's hyperparameters and provides a measure of convergence robust to various Markov chain pathologies. In addition, since GBMs can output predicted class probabilities, we obtain uncertainty measures for $R^*$, which we find provides a useful summary of MCMC convergence. $R^*$ can straightforwardly be incorporated into existing software libraries to provide a complementary convergence metric alongside more established measures.

@@

\section{Method}\label{sec:method}
If Markov chains have not mixed, it is possible to guess (with more accuracy than chance) to which chain a draw belongs from its value. This is possible if there are differences in the sampling distribution for any dimension, $\theta$, in the target distribution (Fig. \ref{fig:marginal}): in this case, if the marginal distributions differ between chains, this information can be used to predict which chain a draw belongs to. It  is also possible to predict the chain that generated a given draw if there are differences in the joint distribution of two (or more) dimensions of the target, even if the marginal distributions are the same (Fig. \ref{fig:joint}).

@@

\subsection{Heterogeneity in chain variance: autoregressive example}\label{sec:heterogeneity}
In this section, we show how $R^*$ is able to detect heterogeneous variance across Markov chains. Apart from demonstrating how $R^*$ performs, this section also investigates the sensitivity of this measure: across different training and testing sets (\S\ref{sec:heterogeneity_stochasticity}) and different draws from the ML model predicted probability simplex (\S\ref{sec:heterogeneity_uncertainty}); to differing numbers of chains (\S\ref{sec:heterogeneity_numchains}); and, finally, to the ML model hyperparameters (\S\ref{sec:heterogeneity_robustness}). The experiments we use to study these issues are all of similar form to the following data generating process: four Markov chains are generated, where each samples from an autoregressive order 1 (AR(1)) process of the form,
%
\begin{equation}\label{eq:ar1}
X_t = \rho X_{t-1} + \epsilon_t,
\end{equation}
%
where $\epsilon_t\stackrel{i.i.d.}{\sim}\text{normal}(0, \sigma)$, $\rho=0.3$ and $t=1,2,...,2000$. Three of the chains share the same $\sigma=1$, whereas the other chain has $\sigma=1/3$, so that it has $1/3$ of the (unconditional) standard deviation of the others.

@@

Rather than run the MCMC sampler for more iterations, we move to a ``non-centered'' parameterisation, which introduces auxillary variables $\boldsymbol{z}\in\mathbb{R}^{250}$ that don't affect $p(\boldsymbol{x})$ but facilitate sampling from it. This model has the form,
%
\begin{align}
\boldsymbol{A}^{-1} = \boldsymbol{L}\boldsymbol{L}^T,\qquad
\boldsymbol{x} = \boldsymbol{L} \boldsymbol{z},\qquad
z_j\sim \text{normal}(0, 1), \text{ for } j = 1,2,...,250.
\end{align}
%
where $\boldsymbol{L}$ is the Cholesky decomposition of the covariance matrix, $\boldsymbol{A}^{-1}$. Fig. \ref{fig:mvt}A shows the $R^*$ distribution resultant from 10,000 NUTS iterations in this case: now the distribution has mean $R^*=1.00$. Fig. \ref{fig:mvt}B shows the $\widehat{R}$ values for each $x$ parameter in this model, and, echoing the result for $R^*$, $\widehat{R}<1.01$ in all cases; further, bulk- and tail-$\text{ESS}>400$ for all parameters.

@@

In addition, we also use Stan to sample from an ``alternative'' parameterisation of the Cauchy, based on a scale mixture of Gaussians \citep{vehtari2019rank},
%
\begin{align}
a_j \sim  \text{normal}(0,1), \qquad
b_j \sim  \text{Gamma}(0.5, 0.5), \qquad
x_j =  a_j/\sqrt{b_j}.
\end{align}
%
The distribution of the $x$ vector is the same under both parameterisations, although the thin-tailed $(a,b)$ vectors define a higher dimensional posterior that improves sampling efficiency.

@@

Comparing our measure of convergence that requires knowing the actual target distribution (\textit{quantile-$R^2$}; in Fig. \ref{fig:cauchy_convergence}A), with the various heuristic measures, all show a similar pattern: as sample size increases, the various statistics tend towards convergence. The rate at which these converge differs though, and $R^*$ (Fig. \ref{fig:cauchy_convergence}D) appears at least, qualitatively, most similar to \textit{quantile-$R^2$}.

@@

The model can be parameterised in two ways, as described in \cite{vehtari2019rank}. The simplest way is referred to as the ``centered'' parameterisation and exactly mirrors the underlying statistical model,
%
\begin{align*}
\theta_j &\sim \text{normal}(\mu, \tau), \\
y_j &\sim \text{normal}(\theta_j, \sigma_j).
\end{align*}
%
The ``non-centered'' parameterisation recodes this model in a way that does not affect the joint distribution of $(\theta, \mu, \tau, \sigma)$ but makes it easier to sample from it, by introducing auxillary variables, $\tilde \theta_j$. This can be written as,
%
\begin{align*}
\tilde{\theta}_j &\sim \text{normal}(0, 1), \\
\theta_j &= \mu + \tau \tilde{\theta}_j,\\
y_j &\sim \text{normal}(\theta_j, \sigma_j).
\end{align*}
%
In both cases, $\theta_j$ are the treatment effects in the eight schools, and $(\mu, \tau)$ represent the population mean and standard deviation 
of the distribution of these effects. In the centered parameterization, the $\theta_j$ are parameters, whereas in the non-centered parameterization, the $\tilde{\theta}_j$ are parameters and $\theta_j$ is a derived quantity.

@@

\subsubsection{Understanding chain classification}
To probe the predictive power of the ML classifier, we investigated how predictive accuracy varies across parameter space. After fitting the GBM model, we group MCMC draws into deciles and draw from the $R^*$ distribution for each decile. In Fig. \ref{fig:eight_schools_r_star_quantiles}, we show the results of this exercise for (A) $\mu$ and (B) $\tau$. In the left-hand column of this figure, we show the path of four MCMC chains (here we did not split chains when calculating $R^*$ to simplify visualisations) across the quantiles of each parameter space. To the right of each trace plot, we show the marginal distributions for each chain. In the right-hand column, we show 40 $R^*$ draws for each decile, which were generated according to Algorithm \ref{alg:R_star_uncertainty} using a GBM fit to all draws. In essence, the left-hand panels explain the variation in $R^*$ in the right-hand panels: if chains become stuck in regions of parameter space, this causes differences between the marginal distributions of the chains; these differences, in turn, allow a ML model to predict the generative chain in those same sticky regions. For example, for $\mu$, the purple chain became stuck around the middle quantile, forcing a difference in its marginal distribution in that region, which resulted in $R^*>1$ for the corresponding decile. Similarly, for $\tau$, the purple chain became stuck in the lowest quantiles, elevating its marginal distribution there and resulting in improved ML predictive accuracy.

@@

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1.0\textwidth]{eight_schools_r_star_quantiles.pdf}}
	\caption{\textbf{Eight schools example: quantile $R^*$ plots.} Row A shows plots for $\mu$; Row B for $\tau$. In each row, we show the path of the four individual chains above and 40 $R^*$ draws obtained using Algorithm \ref{alg:R_star_uncertainty} for each parameter quantile below. To the right of each trace plot, we show the marginal distribution of each chain estimated via kernel density estimation using Gaussian kernels. Note that, in the right-hand plots, jitter has been added to the data points.}
	\label{fig:eight_schools_r_star_quantiles}
\end{figure}

@@

In \S\ref{sec:prostate}, we fit Bayesian models with many 1000s of parameters then used $R^*$ to diagnose convergence, finding that $R^*$ was considerably more expensive to calculate than $\widehat{R}$. The computational cost of training GBMs has been suggested to be $\mathcal{O}(N k)$, where $N$ is the number of training data points and $k$ is the number of dimensions of the target \citep{complexity2018}. If so, this suggests that larger models (usually needing more MCMC iterations) may currently be beyond the reach of $R^*$. That said, it is possible to reduce the runtime for $R^*$ using thinned draws (although this risks losing chain idiosyncracies) and using a subset of dimensions (although this risks losing problematic dimensions). Indeed, in \S\ref{sec:multivariate_normal_250}, \S\ref{sec:cauchy} and \S\ref{sec:prostate}, we use these strategies and, nonetheless, find that $R^*$ provides a stringent measure of convergence.

@@

\section{Acknowledgements}
The authors would like to thank Andrew Gelman and Hanne Oberman for comments on previous drafts of the paper that were submitted to arXiv.
